---
title: "Quasi and Psedo Random Process"
author: "Vincent Lee"
date: "10/4/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Sampling Methods
This project facilitates understanding of generating pseudo-random and quasi-random numbers and application of Monte Carlo method.


### 1. Generate uniformly distributed random numbers
```{r}
#simulate pseudo random numbers uniformly distribution on [0,1]
#pseudo-random process will follow a pattern over time
library(random)
set.seed(100)
Sample <-runif(1000,0,1)


#quasi-random will reduce the gaps between random numbers over time
nFlips <- 1000
dataFromRandom<-randomNumbers(n=nFlips, min=0, max=1, col=1, base=2, check=TRUE)
head(dataFromRandom)

```

###2. Turn binary sequence to uniform random numbers
Turn your sequence of {0,1} into uniform random numbers on [0,1].

Create function that turns a sequence of zeros and ones of length n into decimal form.
```{r}
bitsToInt<-function(x) {
    packBits(rev(c(rep(FALSE, 32-length(x)%%32), as.logical(x))), "integer")
}
bitsToInt(c(1,1,1,1,1,0))
```

Turn the sequence of zeros and ones dataFromRandom of length 1000 into a matrix with 10 columns and 100 rows
```{r}
Binary.matrix<-matrix(dataFromRandom,ncol=10)
head(Binary.matrix)
```

Transform each row of the matrix into decimal format using bin2dec() and divide the numbers by 210 to make real numbers in [0,1].
```{r}
dataFromRandom.dec<-apply(Binary.matrix,1,bitsToInt)/2^10
head(dataFromRandom.dec)
```

###3. Test random number generators

###3.1 Test Sample from runif()

Plot the histogram of "Sample"
```{r}
Sample.histogram<-hist(Sample)
Sample.histogram
```

Test the histogram
```{r}
(Sample.histogram.mean<-mean(Sample.histogram$density))
```

```{r}
(Sample.histogram.sd<-sd(Sample.histogram$density))
```

```{r}
plot(Sample.histogram,freq=FALSE,ylim=c(0,Sample.histogram.mean+2*Sample.histogram.sd))
#h is y value, v is x value
abline(h = Sample.histogram.mean)
abline(h = Sample.histogram.mean+1.96*Sample.histogram.sd,
       col="red",lty=2)
abline(h = Sample.histogram.mean-1.96*Sample.histogram.sd,
       col="red",lty=2)
```

### Moments of the Sample

```{r}
(Sample.histogram.mean<-mean(Sample))
(Sample.histogram.sd<-sd(Sample))
```

### Summary of the Sample

```{r}
summary(Sample)
```

### Testing on quasi-random numbers from dataFromRandom

```{r}
(Sample.histogram.mean<-mean(Sample.histogram$density))
(Sample.histogram.sd<-sd(Sample.histogram$density))
plot(Sample.histogram,freq=FALSE,ylim=c(0,Sample.histogram.mean+2*Sample.histogram.sd))
abline(h=Sample.histogram.mean)
abline(h=Sample.histogram.mean+1.96*Sample.histogram.sd,col="red",lty=2)
abline(h=Sample.histogram.mean-1.96*Sample.histogram.sd,col="red",lty=2)
```

### Moments
Variance for quasi-random numbers is higher than pseudo-random numbers: 

```{r}
(Sample.mean<-mean(dataFromRandom.dec))
(Sample.variance<-var(dataFromRandom.dec))
```

### Test independence of the sequence of zeros and ones

### Turning point test

Turning point test is to test whether sequence of numbers is random or not. 
Null hypothesis is that there is randomness in the number sequence. 
Alaternative hypothesis is serial correlation in the sequence. If test has significantly small p-value (less than 0.05), we reject the null hypothesis. 

Based on findings below, dataFromRandom is indeed not so random as a proof for pseudo-random numbers: 

```{r}
suppressWarnings(library(randtests))
turning.point.test(dataFromRandom.dec)
```


